% Aknowledgement
\vspace*{4cm}
\begin{center}
    {\Large \bf Acknowledgement}
\end{center} \vskip 0.5cm \vskip 0.5cm

I would like to express my deepest gratitude to the authors of the three seminal works that have significantly influenced the direction and depth of my research on "Optimizing LLM Integration Patterns in Enterprise Applications. A Comparative Study of REST, MCP, and Streaming Protocols."

Firstly, I am indebted to Vaswani et al. for their groundbreaking work on the Transformer architecture, as presented in "Attention Is All You Need." Their innovative approach to sequence transduction using attention mechanisms has laid the foundation for modern large language models (LLMs) and inspired the technical underpinnings of this thesis.

Secondly, I extend my heartfelt thanks to the OpenAI team for their pioneering research on GPT-3, as detailed in "Language Models are Few-Shot Learners." Their exploration of scaling LLMs and the introduction of few-shot learning capabilities have provided invaluable insights into the practical deployment and integration of LLMs in enterprise applications.

Lastly, I am grateful to the authors of the DiXtill paper for their novel contributions to knowledge distillation and explainable AI. Their work has been instrumental in shaping my understanding of efficient LLM deployment and the importance of interpretability in enterprise contexts.

This thesis would not have been possible without the inspiration and knowledge derived from these exceptional works. I am deeply appreciative of their contributions to the field of artificial intelligence and their impact on my academic journey.

\bigskip

I would also like to thank my advisors, colleagues, and family for their unwavering support and encouragement throughout this research endeavor.
